---
title: "Terms until Graduation"
author: "Sameer Bhatnagar"
date: "January 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
knitr::read_chunk('predict-science-enrollment.R')
```

```{r load-data}
```

I start with admissions tables to find students admitted to Science, and then find when those students graduated. I then count the terms between their admission and their graduation.

```{r calculate-mean-time-to-graduation-by-cohort, echo=T,message=FALSE}
```

The numbers decrease as we get closer to present, as most Cohort A students have not graudted yet (those we are seeing are program transfers,etc.)


Below I show the distributions throgh box plots. Easily 25% of students from fall cohorts take longer than 4 semesters.
```{r calculate-mean-time-to-graduation-by-cohort-box, echo=T,message=FALSE,warning=FALSE}
```

This is maybe a clearer picture.
```{r calculate-mean-time-to-graduation-by-cohort-bar, echo=T,message=FALSE,warning=FALSE}
```