---
title: "pipeline-v5"
author: "Sameer Bhatnagar"
date: "December 12, 2016"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
knitr::read_chunk('pipeline-v5.R')
```

```{r load-data, echo=FALSE}
```

## Counting Students 
First Let us just make sure we are starting with the right student lists for the different profiles admitted in 20133 and 20141

```{r student-counts}
```

For ESP-DSP, my number matches yours __if we do not__ include people admitted to *09162* and *09164*, however the number is more reasonable if we do.

```{r student-counts-ESPDSP}
```

I continue on from this point keeping the same conventiona as you, with the lower number of students in ESP-DSP.

## Seat Count for All profiles for both Fall and Winter Admits
I give this big table mostly because our numbers _still don't match_ exactly in the next section. Maybe this full table can help figure out why.

```{r seat-counts}
```

## Collapse All Years into one 'Representative Year'
This should match your numbers exactly, and they are still off. I am not sure why. Maybe we can debug using the table above, and my source code (in Appendix at end of this document)

```{r collapse-all-years}
```


## Make projections with additional admits
The fun part: I scale up to 300 admits for Pure & Applied (the way you did), but leave the other profiles alone (unlike your projections). Before I do that as well, we should figure out why the _actual seat counts_ are not exactly the same as yours.

```{r projections-new-admits}
```

## Projections for past years
I go back until 2010, since that is the first year Enviro profile started. I also only look at the bottleneck courses, for readability.

```{r projections-all-years}
```

## Sanity Check
Just to make sure, let's add 100 students to Health Science alone, check if any of said bottleneck courses actually max out.
```{r sanity-check}
```

Looks like  Bio2 in Fall is only the problem?!


